---
title: "Tidy Data"
#author: Talea Cornelius
#date: 09/17/2020
output: github_document
---

```{r setup}
library(tidyverse)
library(haven)

options(tibble.print_min = 5)
```
```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
pulse_data
```
```{r wide to long}
pulse_tidy_data = 
  pivot_longer(
    pulse_data, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    values_to = "bdi")

pulse_tidy_data
```

```{r wide to long}
#remove prefix from names

pulse_tidy_data = 
  pivot_longer(
    pulse_data, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_data
```
```{r pipe data wide to long and clean}

pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>% #str_c treats missing values properly, rather than thinking NA is a string
  arrange(id, visit)

print(pulse_data, n = 12)

```

```{r}
read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  count(group) 
```

```{r separate dose and day from group variable}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(
    dose = str_to_lower(dose),
    wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)

litters_data
```
Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    c(gd0_weight, gd18_weight),
    names_to = "gd",
    values_to = "weight") %>%
  mutate(gd = recode(gd,"gd0_weight" = 0, "gd18_weight" = 18))

select(litters_data, litter_number, gd, weight)

```
```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result
```

```{r}
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```
We’re pretty much there now – in some cases you migth use select to reorder columns, and (depending on your goal) use knitr::kable() to produce a nicer table for reading.

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")

view(fellowship_ring)
view(two_towers)
view(return_king)
```

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "sex", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 

lotr_tidy
```
There are four major ways join dataframes x and y:

Inner: keeps data that appear in both x and y
Left: keeps data that appear in x
Right: keeps data that appear in y
Full: keeps data that appear in either x or y

Left joins are the most common, because they add data from a smaller table y into a larger table x without removing anything from x.

As an example, consider the data tables in FAS_pups.csv and FAS_litters.csv, which are related through the Litter Number variable. The former contains data unique to each pup, and the latter contains data unique to each litter. We can combine these using a left join of litter data into pup data; doing so retains data on each pup and adds data in new columns.

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))

fas_data = 
  left_join(pup_data, litter_data, by = "litter_number")
fas_data
```
Note that joining is not particularly amenable to the %>% operator because it is fundamentally non-linear: two separate datasets are coming together, rather than a single dataset being processed in a step-by-step fashion.


```{r}
surv_os = 
  read_csv("./data/surv_os.csv", col_types = "cc") %>%
  janitor::clean_names() %>%
  rename(uni = what_is_your_uni, os = what_operating_system_do_you_use)

surv_p_git = 
  read_csv("./data/surv_program_git.csv", col_types = "ccc") %>%
  janitor::clean_names() %>%
  rename(uni = what_is_your_uni, program = what_is_your_degree_program,
         experience = which_most_accurately_describes_your_experience_with_git)

surv_data = 
  left_join(surv_os, surv_p_git, by = "uni")
surv_data
```






